On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-01-23 TTau_Halpha/kurosawa.tex
Minor change. (RK)
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@627 25464f21-6099-402c-b30a-640f1d40d38b

The main aim of this paper is to find a simple kinematic model which
can reproduce the wide variety of the observed profiles, and to
perform empirical studies of line formation in an attempt to place
morphological classification schemes on a firmer physical footing. We
will also discuss whether our model is consistent with some
predictions made by recent MHD studies i.e.~
$\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$
(e.g.~\citealt{koenigl:2000}).

@@

@@

    \includegraphics[%
      scale=0.5]{figures/overview3.eps}

@@

  \caption{The basic model configuration. The system consists of four
    components: (1)~the continuum source located at the origin $\left( o
    \right)$ of the cartesian coordinates $\left(x,y,z\right)$ -- the
    $y$-axis is into the paper, (2)~the magnetospheric accretion flow,
    (3)~the bipolar wind outflow, and (4)~the accretion disc. The density
    distribution is symmetric around the $z$-axis. The innermost radius of
    the magnetosphere (at the equatorial plane) coincides with the the
    inner radius of the accretion disc. From the innermost part of the
    accretion disc, the gas falls freely, moving along the magnetic field
    onto the surface of the star. The bipolar wind starts from just
    outside the largest radius 
    ($R_{\mathrm{mo}}$) of the magnetosphere (dotted line). }

@@

\begin{tabular}{lccccccccccc}
\hline 
Parameters&
$R_{*}$&
$M_{*}$&
$T_{\mathrm{ph}}$&
$R_{\mathrm{mi}}$&
$R_{\mathrm{mo}}$&
$\dot{M}_{\mathrm{acc}}$&
$\dot{M}_{\mathrm{wind}}$&
$v_{\infty}$&
$b$&
$R_{\mathrm{di}}$&
$R_{\mathrm{do}}$\tabularnewline
&
$\left[R_{\sun}\right]$&
$\left[M_{\sun}\right]$&
$\left[\mathrm{K}\right]$&
$\left[R_{*}\right]$&
$\left[R_{*}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$[\mathrm{km\, s^{-1}}]$&
$[-]$&
$\left[R_{*}\right]$&
$\left[\mathrm{au}\right]$\tabularnewline
\hline
Standard&
$2.0$&
$0.5$&
$4000$&
$2.2$&
$3.0$&
$10^{-7}$&
$10^{-8}$&
$210$&
$4.0$&
$2.2$&
$100$\tabularnewline
\hline
\end{tabular}

@@

The magneto-centrifugal wind paradigm, first proposed by
\citet{blandford:1982}, has been often used to model the large-scale
wind structure of T Tauri stars, or the observed optical jets
(e.g.~HH~30 jet by \citealt{burrows:1996}; \citealt{ray:1996}). The
launching of the wind from a Keplerian disc is typically done by
treating the equatorial plane of the disc as a mass-injecting boundary
condition (e.g., \citealt{shu:1994}; \citealt{ustyugova:1995};
\citealt{ouyed:1997}; \citealt{krasnopolsky:2003}).  Depending on the
location of the open magnetic fields anchored to the disc, two
different types of winds are produced. If the field is constrained to
be near the co-rotation radius of stellar magnetosphere, an
{}``X-wind'' \citep{shu:1994} is produces. If the open field lines are
located in a wider area of the disc, a {}``disc-wind'' similar to that
of \citet{koenigl:2000} is produced \citep{krasnopolsky:2003}.  Recent
reviews on the jet/wind-disc connection can be found in
\citet{koenigl:2000} and \citet{pudritz:2005}.

Clearly there are several alternative outflow scenarios. Here we adopt
two simple kinematical models which broadly represent the results of
MHD simulations. The first is a bipolar wind model, in which the
outflow is collimated along the rotational axes (a model similar to
that of \citet{krasnopolsky:2003}). The second model is based on the
disc-wind paradigm, in which the density of the outflow increases
towards the equatorial plane. The difference in latitudinal density
structure between the two outflow models should provide an indication
as to whether line profiles alone can be used to constrain the outflow
geometry. In the following sections we present details of the two
outflow structures.

@@

In a similar fashion to the simple model of \citet{appenzeller:2005},
the following parametrisation of collimated bipolar wind (region 3 in
Fig.~\ref{fig:config_dipolar}) is adopted. The wind velocity field,
$\boldsymbol{v}_{\mathrm{wind}}$, consists of radial and azimuthal
components which depend on the spherical coordinates $r$ and
$\theta$. The radial component $v_{r}\left(r\right)$ is assumed to
follow the classical beta-velocity law \citep[c.f.][]{castor:1979},
and the azimuthal component $v_{\phi}$ is assumed to be a constant
fraction ($\gamma$) of the Keplerian velocity for a given distance
($w=\sqrt{x^{2}+y^{2}}$) from the symmetry axis ($z$-axis) , i.e.:
\begin{equation}
     \boldsymbol{v}_{\mathrm{wind}}  = 
     v_{r}\hat{\boldsymbol{r}}+v_{\phi}\hat{\boldsymbol{\bphi}}
  \label{eq:wind_velocity_def}
\end{equation}
where 
\begin{equation}
   v_{r}\left(r\right) =
   v_{r0}+\left(v_{\infty}-v_{r0}\right)\left(1-\frac{R_{\mathrm{mo}}}{r}\right)^{\beta}\,,
   \label{eq:wind_velocity_radial_part}
\end{equation}
and
\begin{equation}
      v_{\phi}\left(w\right)=\gamma\,\left(\frac{GM_{*}}{w}\right)^{1/2}\,.
   \label{eq:wind_velocity_azimuth_part}
\end{equation}
Note that the base of the wind starts at $r=R_{\mathrm{mo}}$. This
naturally leads to a density discontinuity, but allows a simple
parameterisation of the wind whereas a self-consistent interface
between the magnetosphere and wind would require a full MHD
calculation that is beyond the scope of this paper. The range of polar
angle for the wind is restricted to
$\theta>\left|\theta_{\mathrm{disc}}\right|$ where
$\theta_{\mathrm{disc}}$ is the opening angle of the accretion
disc. $v_{r0}$ is the small radial velocity at the base of the wind
($r=R_{\mathrm{mo}}$). Normally, $v_{r0}=10\,\mathrm{km\, s^{-1}}$
which is approximately equal to the thermal velocity of hydrogen with
$T=7500\,\mathrm{K}$. Following \citet{appenzeller:2005},
$\gamma=0.05$ is adopted. The dependency of $v_{r}$ in polar direction
($\theta=0$) on the values of wind acceleration parameter $\beta$ is
shown in Figure~\ref{fig:wind_vr_rho}. All other parameters describing
the wind are fixed as the standard values given in
Table~\ref{tab:std_parameters}.

The density of the wind is assumed to be axi-symmetric and
separable in $r$ and $\theta$, 
\begin{equation}
    \rho\left(r,\theta\right)=P\left(r\right)\, F\left(\theta\right)
    \label{eq:wind_density_def}
\end{equation}
with 
\begin{equation}
    F\left(\theta\right)=n\,\cos^{b}\theta
    \label{eq:wind_density_theta}
\end{equation}
where $b$ is normally a positive even number (for density symmetric
about the equatorial plane), and $n$ is the angular normalisation
constant. For $b=0$, the wind is spherically symmetric except for
the parts disrupted by the accretion disc. The larger the value of
$b$, the higher the degree of the collimation. By integrating equation~\ref{eq:wind_density_theta}
over angles and normalising the integral to $4\pi$, one finds 
\begin{equation}
  n=\frac{1+b}{1-\cos^{1+b}\theta_{\mathrm{wind}}}\,.
  \label{eq:wind_density_theta_norm}
\end{equation}
Assuming the total mass-loss rate via the wind is
$\dot{M}_{\mathrm{wind}}$ and the mass-flux is constant, the radial
part of the density function is reduced to
$P\left(r\right)=\dot{M}_{\mathrm{wind}}\,\left[4\pi
r^{2}v_{r}\left(r\right)\right]^{-1}$;
hence, equation~\ref{eq:wind_density_def} becomes 
\begin{equation}
  \rho\left(r,\theta\right) = 
  \frac{n\,\dot{M}_{\mathrm{wind}}\,\cos^{b}\theta}{4\pi r^{2}\,
    v_{r}\left(r\right)}\,.
  \label{eq:wind_density_final}
\end{equation}
For a given mass-accretion rate, the wind mass-loss rate in our
typical model is assigned from the ratio of mass-loss to
mass-accretion rate
$\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$, a value
indicated by both observations and MHD calculations
(e.g. \citealt{koenigl:2000}).  Figure~\ref{fig:wind_vr_rho} shows the
density along a streamline in the polar direction ($\theta=0$), for
different values of $\beta$ with all other parameters fixed at the
standard values (Table~\ref{tab:std_parameters}).  The density is
relatively sensitive to the value of $\beta$ for $r<10\,
R_{\mathrm{mo}}$.

@@

Table~\ref{tab:ha_ew_acc_models} shows the EW for the models in
Figure~\ref{fig:atlas_acc_models}.  The EWs for half the models
fall within the range of EWs ($\sim 3$~\AA\, to $\sim 160$~\AA)
measured by \citet{alencar:2000}. The equivalent widths of models 
with the lowest mass accretion rates fall below the minimum EW
observed by  \citet{alencar:2000}, while several models would
be designated as weak-lined T~Tauri stars using the traditional
10\,\AA\ cut-off.

xxxx might want to talk about line widths as an accretion 
indicator here as well (white and basri).

The dependency of the line profile on inclination angle ($i$) is
demonstrated in Figure~\ref{fig:acc_inclination_effect}. The model
has  $T_{\mathrm{max}}=7500\,\mathrm{K}$ and
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$).  The
figure shows that the peak (normalised) flux decreases as the
inclination angle increases, as does the EW.

xxxx check reasons for inclination dependence xxxx

Because of the geometry of the
magnetospheric accretion (c.f. Figure~\ref{fig:config_dipolar}) and of
the presence of the gas with the highest velocity close to the stellar
surface, the highest red-shifted line-of-sight velocity is visible
only at the high inclination angles. This explains the wider
appearance of the profile with $i=80^{\circ}$compared to the
relatively narrow line appearance of the profile with
$i=10^{\circ}$.

A similar EW dependence on
inclination angle is found for the models with different 
magnetospheric temperatures ($T_{\mathrm{max}}=6500$,
$8500$, $9500\,\mathrm{K}$).

@@

    \includegraphics[%
      clip,
      scale=0.45]{figures/broadening.eps}

@@

    \includegraphics[%
      clip,
      scale=0.45]{figures/acc_inc_effect.eps}

@@

\caption{Summary of H$\alpha$ equivalent widths from the
magnetospheric accretion flow models shown in
Figure~\ref{fig:atlas_acc_models}.}

@@

@@

  \caption{H$\alpha$ model profiles computed with wide ranges of the
    isothermal wind temperature ($T_{\mathrm{wind}}$) and the wind
    acceleration rate ($\beta$). The profiles are computed with the wind
    component only (no magnetospheric accretion flow or accretion
    disc). The `standard' model parameters (Table~\ref{tab:std_parameters}) are
    applied. The wind mass-loss rate ($\dot{M}_{\mathrm{wind}}$),
    the inclination ($i$) and the degree of columniation ($b$) are
    fixed at $10^{-8}~{\mathrm{M_{\sun}~yr^{-1}}}$, $55^{\circ}$ and
    $4.0$\, respectively. The temperature (indicated along the vertical
    axis) of the model increases from top to bottom, and the wind
    acceleration rate increases from left to right. The P~Cygni profile
    are seen for lower $T_{\mathrm{wind}}$ and lower $\beta$ (slow
    acceleration) models.  The position of the absorption the component
    moves toward the line centre as $\beta$ increases. The effect of
    line broadening becomes more prominent for higher $T_\mathrm{wind}$
    and higher $\beta$ models.}

@@

Firstly, we consider the disc-wind models without the magnetosphere.
Figure~\ref{fig:atlas_discwind_only} shows the H$\alpha$ profiles
computed using only the disc-wind component (c.f. Figure~\ref{fig:config_discwind})
for the same isothermal disc-wind temperature ($T_{\mathrm{dw}}$)
and the wind acceleration parameter ($\beta$) ranges used for the
bipolar wind models (c.f. xxxx cite bipolar figure here). The parameters used for the central star are
as in section~\ref{sub:Continuum-Source}, and the disc-wind
parameters are summarised in Table~\ref{tab:std_discwind_parameters}.
The figure shows that the wind emission increases as $\beta$ and $T_{\mathrm{dw}}$
increase. The $\beta$ dependence arises simply because for large $\beta$
there is a greater mass of wind material close to the star.
As with  the bipolar wind models, the location of the
P-Cyg absorption component moves toward the line centre as $\beta$
increases.

The morphology of the profile exhibited by the model changes
from FU-Ori type for the models with small $\beta$ to the stronger
emission types which are more commonly seen in the observed H$\alpha$
(c.f.~\citealt{reipurth:1996}) as the value of $\beta$
increases.  We find that the profile shapes
are relatively insensitive to the value of the wind collimation parameter
$d$. The dependency on the disc wind mass-loss rate of the profile
shape is similar to that of the bipolar wind models
(Figure~\ref{fig:effect_mdot_disc_wind_acc}).

@@

As in the disc-wind only models
(Figure~\ref{fig:atlas_discwind_only}), the location of the absorption
component moves toward the line centre as the value of $\beta$
increases.  The figure shows that in
the models with smaller $\beta$ and $T_{\mathrm{dw}}$, the line
emission from the magnetosphere dominates, while this situation reverses for the
models with larger $\beta$ and $T_{\mathrm{dw}}$. The model with
$\left(T_{\mathrm{dw}},\,\beta\right)=\left(9000\,\mathrm{K},\,2.0\right)$
results in an H$\alpha$ profile that is far too strong to be compatible
 with observations (\citet{reipurth:1996} and \citet{alencar:2000}).

Although the models in Figure~\ref{fig:atlas_discwind_acc_model} are
computed with limited ranges of  $T_{\mathrm{dw}}$ and $\beta$, the
resulting profiles exhibit a wide variety of line profile shapes,
many of  which are similar to the types of H$\alpha$ profiles seen in the
observations (c.f.~\citealt{reipurth:1996}).   The shape of the
primary and secondary peaks in the profiles are slightly {}``rounder''
in comparison to those of the accretion, bipolar wind, disc hybrid model
shown in Figure~\ref{fig:atlas_disc_wind_acc_model}.  The rounder
peaks are the result of wind rotation.

xxxx think of a way of talking about rotation in comparison between
the disc wind model and the bipolar wind model xxxx

An examination of the profiles in Figures~\ref{fig:atlas_discwind_only} and
Figure~\ref{fig:atlas_discwind_acc_model} reveals some of problems
with uniqueness that must inevitably occur when trying to assign
an individual profile as wind or accretion dominated i.e. very similar profiles may arise
from very different circumstellar geometries. For example, the disc-wind
only model with $\left(T_{\mathrm{wind}},\,\beta\right)=\left(8000\,\mathrm{K},\,1.0\right)$
(Figures~\ref{fig:atlas_discwind_only}) and the  accretion dominated model with
 $\left(T_{\mathrm{wind}},\,\beta\right)=\left(9000\,\mathrm{K},\,0.5\right)$
in (Figure~\ref{fig:atlas_discwind_acc_model}) give very similar profiles.

@@

\caption{A summary of the standard disc-wind model parameters used
in section~\ref{sub:result-discwind-knigge}. See also
section~\ref{sub:Alternative-model-disc-wind}.}

@@

The inclination dependency of the line equivalent width predicted by
the bipolar wind model agree with trends seen in the observation, but
the disc-wind model does not. With the standard magnetospheric
configuration (Table~\ref{tab:std_parameters}), the ratio of the
mass-loss to mass-accretion rates ($\mu$) used in the bipolar wind was
set to be 4 --10 times smaller than that predicated by
magneto-hydrodynamical (MHD) calculations (c.f.~$\mu\sim0.1;$
\citealt{koenigl:2000}) to avoid the blue-shifted absorption component
becoming unrealistically too strong; however $\mu$ used in the
disc-wind models are consistent with MHD calculations.

>>>>

The main aim of this paper is to find a simple kinematic model which
can reproduce the wide variety of the observed profiles, and to
perform empirical studies of line formation in an attempt to place
morphological classification schemes on a firmer physical footing. We
will also discuss whether our model is consistent with some
predictions made by recent (magneto-hydrodynamics) MHD studies i.e.~
$\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$
(e.g.~\citealt{koenigl:2000}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \begin{figure*}

%%   \begin{center}

%%     \includegraphics[%
%%       scale=0.5]{figures/overview3.eps}

%%   \end{center}

%%   \caption{The basic model configuration. The system consists of four
%%     components: (1)~the continuum source located at the origin $\left( o
%%     \right)$ of the cartesian coordinates $\left(x,y,z\right)$ -- the
%%     $y$-axis is into the paper, (2)~the magnetospheric accretion flow,
%%     (3)~the bipolar wind outflow, and (4)~the accretion disc. The density
%%     distribution is symmetric around the $z$-axis. The innermost radius of
%%     the magnetosphere (at the equatorial plane) coincides with the the
%%     inner radius of the accretion disc. From the innermost part of the
%%     accretion disc, the gas falls freely, moving along the magnetic field
%%     onto the surface of the star. The bipolar wind starts from just
%%     outside the largest radius 
%%     ($R_{\mathrm{mo}}$) of the magnetosphere (dotted line). }

%%   \label{fig:config_dipolar}

%% \end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}

@@

    \includegraphics[%
      scale=0.3]{figures/overview3_alt.eps}

@@

  \caption{The basic model configuration. The system consists of four
    components: (1)~the continuum source located at the origin $\left( o
    \right)$ of the cartesian coordinates $\left(x,y,z\right)$ -- the
    $y$-axis is into the paper, (2)~the magnetospheric accretion flow,
    (3)~the bipolar wind outflow, and (4)~the accretion disc. The density
    distribution is symmetric around the $z$-axis. The innermost radius of
    the magnetosphere (at the equatorial plane) coincides with the the
    inner radius of the accretion disc. From the innermost part of the
    accretion disc, the gas falls freely, moving along the magnetic field
    onto the surface of the star. The wind density is enhaced toward
    the polar directions to mimic the collimated wind structure of the
    magneto-hydrodynamical simulations (e.g.~\citealt{krasnopolsky:2003}). }

@@

\begin{tabular}{lccccccccccc}
\hline 
Parameters&
$R_{*}$&
$M_{*}$&
$T_{\mathrm{ph}}$&
$R_{\mathrm{mi}}$&
$R_{\mathrm{mo}}$&
$\dot{M}_{\mathrm{acc}}$&
$\dot{M}_{\mathrm{wind}}$&
$v_{\infty}$&
$b$&
$R_{\mathrm{di}}$&
$R_{\mathrm{do}}$\tabularnewline
&
$\left[R_{\sun}\right]$&
$\left[M_{\sun}\right]$&
$\left[\mathrm{K}\right]$&
$\left[R_{*}\right]$&
$\left[R_{*}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$[\mathrm{km\, s^{-1}}]$&
$[-]$&
$\left[R_{*}\right]$&
$\left[\mathrm{au}\right]$\tabularnewline
\hline
&
$2.0$&
$0.5$&
$4000$&
$2.2$&
$3.0$&
$10^{-7}$&
$10^{-8}$&
$210$&
$4.0$&
$2.2$&
$100$\tabularnewline
\hline
\end{tabular}

@@

The magneto-centrifugal wind paradigm, first proposed by
\citet{blandford:1982}, has been often used to model the large-scale
wind structure of T Tauri stars, or the observed optical jets
(e.g.~HH~30 jet by \citealt{burrows:1996}; \citealt{ray:1996}). The
launching of the wind from a Keplerian disc is typically done by
treating the equatorial plane of the disc as a mass-injecting boundary
condition (e.g., \citealt{shu:1994}; \citealt{ustyugova:1995};
\citealt{ouyed:1997}; \citealt{krasnopolsky:2003}).  Depending on the
location of the open magnetic fields anchored to the disc, two
different types of winds are produced. If the field is constrained to
be near the co-rotation radius of stellar magnetosphere, an
{}``X-wind'' \citep{shu:1994} is produced. If the open field lines are
located in a wider area of the disc, a {}``disc-wind'' similar to that
of \citet{koenigl:2000} is produced \citep{krasnopolsky:2003}.  Recent
reviews on the jet/wind-disc connection can be found in
\citet{koenigl:2000} and \citet{pudritz:2005}.

Clearly there are several alternative outflow scenarios. Here we adopt
two simple kinematical models which broadly represent the results of
MHD simulations. The first is a bipolar wind model, in which the
outflow is collimated along the rotational axes (a model similar to
that of \citealt{krasnopolsky:2003} in large scale, $>10$~au). The
second model is based on the disc-wind paradigm, in which the density
of the outflow increases towards the equatorial plane. The difference
in latitudinal density structure between the two outflow models should
provide an indication as to whether line profiles alone can be used to
constrain the outflow geometry. In the following sections we present
details of the two outflow structures.

@@

In a similar fashion to the simple model of \citet{appenzeller:2005},
the following parametrisation of collimated bipolar wind (region 3 in
Fig.~\ref{fig:config_dipolar}) is adopted. The wind velocity field,
$\boldsymbol{v}_{\mathrm{wind}}$, consists of radial and azimuthal
components which depend on the spherical coordinates $r$ and
$\theta$. The radial component $v_{r}\left(r\right)$ is assumed to
follow the classical beta-velocity law \citep[c.f.][]{castor:1979},
and the azimuthal component $v_{\phi}$ is assumed to be a constant
fraction ($\gamma$) of the Keplerian velocity for a given distance
($w=\sqrt{x^{2}+y^{2}}$) from the symmetry axis ($z$-axis) , i.e.:
\begin{equation}
     \boldsymbol{v}_{\mathrm{wind}}  = 
     v_{r}\hat{\boldsymbol{r}}+v_{\phi}\hat{\boldsymbol{\bphi}}
  \label{eq:wind_velocity_def}
\end{equation}
where 
\begin{equation}
   v_{r}\left(r\right) =
   v_{r0}+\left(v_{\infty}-v_{r0}\right)\left(1-\frac{R_{\mathrm{mo}}}{r}\right)^{\beta}\,,
   \label{eq:wind_velocity_radial_part}
\end{equation}
and
\begin{equation}
      v_{\phi}\left(w\right)=\gamma\,\left(\frac{GM_{*}}{w}\right)^{1/2}\,.
   \label{eq:wind_velocity_azimuth_part}
\end{equation}
Note that the base of the wind starts at $r=R_{\mathrm{mo}}$. This
naturally leads to a density discontinuity, but allows a simple
parameterisation of the wind whereas a self-consistent interface
between the magnetosphere and wind would require a full MHD
calculation that is beyond the scope of this paper. The range of polar
angle for the wind is restricted to
$\theta>\left|\theta_{\mathrm{disc}}\right|$ where
$\theta_{\mathrm{disc}}$ is the opening angle of the accretion
disc. $v_{r0}$ is the small radial velocity at the base of the wind
($r=R_{\mathrm{mo}}$). Normally, $v_{r0}=10\,\mathrm{km\, s^{-1}}$
which is approximately equal to the thermal velocity of hydrogen with
$T=7500\,\mathrm{K}$. Following \citet{appenzeller:2005},
$\gamma=0.05$ is adopted. The dependency of $v_{r}$ in polar direction
($\theta=0$) on the values of wind acceleration parameter $\beta$ is
shown in Figure~\ref{fig:wind_vr_rho}. All other parameters describing
the wind are fixed as the reference values given in
Table~\ref{tab:std_parameters}.

The density of the wind is assumed to be axi-symmetric and
separable in $r$ and $\theta$, 
\begin{equation}
    \rho\left(r,\theta\right)=P\left(r\right)\, F\left(\theta\right)
    \label{eq:wind_density_def}
\end{equation}
with 
\begin{equation}
    F\left(\theta\right)=n\,\cos^{b}\theta
    \label{eq:wind_density_theta}
\end{equation}
where $b$ is normally a positive even number (for density symmetric
about the equatorial plane), and $n$ is the angular normalisation
constant. For $b=0$, the wind is spherically symmetric except for
the parts disrupted by the accretion disc. The larger the value of
$b$, the higher the degree of the collimation. By integrating equation~\ref{eq:wind_density_theta}
over angles and normalising the integral to $4\pi$, one finds 
\begin{equation}
  n=\frac{1+b}{1-\cos^{1+b}\theta_{\mathrm{wind}}}\,.
  \label{eq:wind_density_theta_norm}
\end{equation}
Assuming the total mass-loss rate via the wind is
$\dot{M}_{\mathrm{wind}}$ and the mass-flux is constant, the radial
part of the density function is reduced to
$P\left(r\right)=\dot{M}_{\mathrm{wind}}\,\left[4\pi
r^{2}v_{r}\left(r\right)\right]^{-1}$;
hence, equation~\ref{eq:wind_density_def} becomes 
\begin{equation}
  \rho\left(r,\theta\right) = 
  \frac{n\,\dot{M}_{\mathrm{wind}}\,\cos^{b}\theta}{4\pi r^{2}\,
    v_{r}\left(r\right)}\,.
  \label{eq:wind_density_final}
\end{equation}
For a given mass-accretion rate, the wind mass-loss rate in our
typical model is assigned from the ratio of mass-loss to
mass-accretion rate
$\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$, a value
indicated by both observations and MHD calculations
(e.g. \citealt{koenigl:2000}).  Figure~\ref{fig:wind_vr_rho} shows the
density along a streamline in the polar direction ($\theta=0$), for
different values of $\beta$ with all other parameters fixed at the
reference values (Table~\ref{tab:std_parameters}).  The density is
relatively sensitive to the value of $\beta$ for $r<10\,
R_{\mathrm{mo}}$.

@@

Table~\ref{tab:ha_ew_acc_models} shows the equivalent width (EW) for
the models in Figure~\ref{fig:atlas_acc_models}.  The EWs for half the
models fall within the range of EWs ($\sim 3$~\AA\, to $\sim 160$~\AA)
measured by \citet{alencar:2000}. The EWs of models 
with the lowest mass accretion rates fall below the minimum EW
observed by  \citet{alencar:2000}, while several models would
be designated as weak-lined T~Tauri stars (WTTS) using the traditional
10\,\AA\ cut-off \citep[e.g.][]{herbig:1988}.

xxxx Changed xxxxx

\citet{white:2003} empirically showed that the full width of H$\alpha$ at 10
per~cent of its peak flux (10 per~cent width) is a better
indicator for an accretion than the EW criteria. They proposed that a
T~Tauri star shows the 10 per~cent width greater than
$270\,\mathrm{km\,s^{-1}}$, the star should be a CTTS. The 10 per~cent
widths of the model H$\alpha$ profiles
(Figure~\ref{fig:atlas_acc_models}) are summarised in
Table~\ref{tab:ha_ten_percent_acc_models}.  Using the criteria of
\citet{white:2003}, the half of the models shown in the figure can be
classified as CTTS, and the other half as WTTS.

The dependency of the line profile on inclination angle ($i$) is
demonstrated in Figure~\ref{fig:acc_inclination_effect}. The model
has  $T_{\mathrm{max}}=7500\,\mathrm{K}$ and
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$).  The
figure shows that the peak (normalised) flux decreases as the
inclination angle increases, as does the EW.  The line flux and the
EWs decreases since the fraction of the accretion stream blocked by
the photosphere increases as the inclination increase.

xxxx Changed xxxxx

Because of the geometry of the
magnetospheric accretion (c.f. Figure~\ref{fig:config_dipolar}) and of
the presence of the gas with the highest velocity close to the stellar
surface, the highest red-shifted line-of-sight velocity is visible
only at the high inclination angles. This explains the wider
appearance of the profile with $i=80^{\circ}$compared to the
relatively narrow line appearance of the profile with
$i=10^{\circ}$.  A similar EW dependence on
inclination angle is found for the models with different 
magnetospheric temperatures and the mass-accretion rates.

@@

    \includegraphics[%
      clip,
      scale=0.59]{figures/broadening.eps}

@@

    \includegraphics[%
      clip,
      scale=0.59]{figures/acc_inc_effect.eps}

@@

\caption{Summary of H$\alpha$ equivalent widths (\AA) from the
magnetospheric accretion flow models shown in
Figure~\ref{fig:atlas_acc_models}. Note that the equivalent widths are
positive when the lines are in emission.}

@@

\caption{Summary of 10 per~cent widths ($\mathrm{km\,s^{-1}}$) from the
magnetospheric accretion flow models shown in
Figure~\ref{fig:atlas_acc_models}.}

\label{tab:ha_ten_percent_acc_models}

\begin{tabular}{cccc}
\hline 
$T_{\mathrm{max}}\,\,\left(\mathrm{K}\right)$&
&
$\dot{M}_{\mathrm{acc}}$ ~$\left(\mathrm{M_{\sun}\, yr^{-1}}\right)$&
\tabularnewline
&
$10^{-7}$&
$10^{-8}$&
$10^{-9}$\tabularnewline
\hline 
$6500$&
$300$&
$120$&
$100$\tabularnewline
$7500$&
$360$&
$145$&
$120$\tabularnewline
$8500$&
$790$&
$330$&
$130$\tabularnewline
$9500$&
$1530$&
$520$&
$200$\tabularnewline
\hline
\end{tabular}

@@

  \caption{H$\alpha$ model profiles computed with wide ranges of the
    isothermal wind temperature ($T_{\mathrm{wind}}$) and the wind
    acceleration rate ($\beta$). The profiles are computed with the wind
    component only (no magnetospheric accretion flow or accretion
    disc). The `reference' model parameters (Table~\ref{tab:std_parameters}) are
    applied. The wind mass-loss rate ($\dot{M}_{\mathrm{wind}}$),
    the inclination ($i$) and the degree of columniation ($b$) are
    fixed at $10^{-8}~{\mathrm{M_{\sun}~yr^{-1}}}$, $55^{\circ}$ and
    $4.0$\, respectively. The temperature (indicated along the vertical
    axis) of the model increases from top to bottom, and the wind
    acceleration rate increases from left to right. The P~Cygni profile
    are seen for lower $T_{\mathrm{wind}}$ and lower $\beta$ (slow
    acceleration) models.  The position of the absorption the component
    moves toward the line centre as $\beta$ increases. The effect of
    line broadening becomes more prominent for higher $T_\mathrm{wind}$
    and higher $\beta$ models.}

@@

Firstly, we consider the disc-wind models without the magnetosphere.
Figure~\ref{fig:atlas_discwind_only} shows the H$\alpha$ profiles
computed using only the disc-wind component
(c.f. Figure~\ref{fig:config_discwind}) for the same isothermal
disc-wind temperature ($T_{\mathrm{dw}}$) and the wind acceleration
parameter ($\beta$) ranges used for the bipolar wind models
(c.f. Figures~\ref{fig:atlas_wind_model} and
\ref{fig:atlas_disc_wind_acc_model}).  The parameters used for the
central star are as in section~\ref{sub:Continuum-Source}, and the
disc-wind parameters are summarised in
Table~\ref{tab:std_discwind_parameters}.  The figure shows that the
wind emission increases as $\beta$ and $T_{\mathrm{dw}}$ increase. The
$\beta$ dependence arises simply because for large $\beta$ there is a
greater mass of wind material close to the star.  As with the bipolar
wind models, the location of the P-Cyg absorption component moves
toward the line centre as $\beta$ increases.

The morphology of the profile exhibited by the model changes from
FU-Ori type (c.f.~\citealt{reipurth:1996}) for the models with small
$\beta$ to the stronger emission types which are more commonly seen in
the observed H$\alpha$ (c.f.~\citealt{reipurth:1996}) as the value of
$\beta$ increases.  We find that the profile shapes are relatively
insensitive to the value of the wind collimation parameter $d$. The
dependency on the disc wind mass-loss rate of the profile shape is
similar to that of the bipolar wind models
(Figure~\ref{fig:effect_mdot_disc_wind_acc}).

@@

As in the disc-wind only models
(Figure~\ref{fig:atlas_discwind_only}), the location of the absorption
component moves toward the line centre as the value of $\beta$
increases.  The figure shows that in
the models with smaller $\beta$ and $T_{\mathrm{dw}}$, the line
emission from the magnetosphere dominates, while this situation reverses for the
models with larger $\beta$ and $T_{\mathrm{dw}}$. The model with
$\left(T_{\mathrm{dw}},\,\beta\right)=\left(9000\,\mathrm{K},\,2.0\right)$
results in an H$\alpha$ profile that is far too strong to be compatible
 with observations (\citet{reipurth:1996} and \citet{alencar:2000});
 however, the lower mass-loss rate models (while keeping $\mu$ constant)
 produce the profiles with the line strengths comparable to observations.
Although the models in Figure~\ref{fig:atlas_discwind_acc_model} are
computed with limited ranges of  $T_{\mathrm{dw}}$ and $\beta$, the
resulting profiles exhibit a wide variety of line profile shapes,
many of  which are similar to the types of H$\alpha$ profiles seen in the
observations (c.f.~\citealt{reipurth:1996}).

An examination of the profiles in Figures~\ref{fig:atlas_discwind_only} and
Figure~\ref{fig:atlas_discwind_acc_model} reveals some of problems
with uniqueness that must inevitably occur when trying to assign
an individual profile as wind or accretion dominated i.e. very similar profiles may arise
from very different circumstellar geometries. For example, the disc-wind
only model with $\left(T_{\mathrm{wind}},\,\beta\right)=\left(8000\,\mathrm{K},\,1.0\right)$
 (in Figures~\ref{fig:atlas_discwind_only}) and the  accretion dominated model with
 $\left(T_{\mathrm{wind}},\,\beta\right)=\left(9000\,\mathrm{K},\,0.5\right)$
(in Figure~\ref{fig:atlas_discwind_acc_model}) give very similar profiles.

@@

\caption{A summary of the reference disc-wind model parameters used
in section~\ref{sub:result-discwind-knigge}. See also
section~\ref{sub:Alternative-model-disc-wind}.}

@@

The inclination dependency of the line equivalent width predicted by
the bipolar wind model agree with trends seen in the observation, but
the disc-wind model does not. With the reference magnetospheric
configuration (Table~\ref{tab:std_parameters}), the ratio of the
mass-loss to mass-accretion rates ($\mu$) used in the bipolar wind was
set to be 4 --10 times smaller than that predicated by
magneto-hydrodynamical (MHD) calculations (c.f.~$\mu\sim0.1;$
\citealt{koenigl:2000}) to avoid the blue-shifted absorption component
becoming unrealistically too strong; however $\mu$ used in the
disc-wind models are consistent with MHD calculations.
